---
title: "Whole data visualization"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r Working directory, echo=TRUE, message=FALSE, warning=FALSE}
#setwd("../PIB_workflowr/data")
getwd()
```

```{r Library load, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(Rtsne)
library(reshape2)
library(ggpubr)
library(gghighlight)
library(plotly)
library(corrplot)
library(tibble)
library(tidyr)
library(dplyr)
theme_set(theme_minimal())
```

# Data Loading

```{r}
x <- read_tsv("filenames.tsv")

x$fullpath <- paste("", x$filename, sep="")
r <- list()

for (i in 1:nrow(x)) {

  tmp <- read_tsv(x$fullpath[i], col_names = F, col_types = "dd")

  names(tmp) <- c("mz", "intensity")

  tmp <- tmp %>% bind_cols(x[i,])

  r[[length(r)+1]] <- tmp

}

df <- bind_rows(r)
df$pen<-as.character(df$pen)
str(df)

write_rds(df, file = "full_data.rds")

rm(r,tmp,x)
```

## Forth Square root transform

# How does the data look like?

## Looking at the dataframe

```{r}
head(df,7)

# 4th root transformation of the data
df$intensity<-df$intensity^(1/4)

head(df,7)
```

## Plotting

Basic plot of the three sets of data we have:

```{r}
df %>% ggplot(aes(x=mz,y=intensity))+geom_point(aes(color=pen),alpha=.4) + facet_wrap(~group)
```

Let's check only test and train and facet_wrapping:

```{r}
df %>% filter(group==c("training","test")) %>% ggplot(aes(x=mz,y=intensity))+ geom_point(aes(color=group), size=2, alpha=.4) + facet_wrap(~pen)

df %>% filter(group==c("training","test")) %>% ggplot(aes(x=mz,y=intensity))+ geom_point(aes(color=pen, shape = group), size=2, alpha=.4) + facet_wrap(~group+pen)
```

# Whole data PCA analysis

2d PCA

```{r}
pca_df<-spread(df,mz,intensity)
pr.out <- prcomp(pca_df[,6:(ncol(pca_df)-1)], scale. = TRUE)

autoplot(pr.out, data = pca_df, colour = 'group', shape="group", size=2)+scale_shape_manual(values=c(1, 8, 12))+ scale_color_manual(values=c("black", "darkorange", "dodgerblue"))+geom_text(aes(label=pen),hjust=-1, vjust=1)
```

3d PCA

```{r}

dat_3d <- pr.out$x[,1:3] %>% as_tibble() %>% mutate(group = pca_df$group, pen=pca_df$pen)

fig <- plot_ly(dat_3d, x = ~PC1, y = ~PC2, z = ~PC3, 
               color = ~group,colors = c("black", "darkorange", "dodgerblue"), size = I(100))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'PC1'),
                                   yaxis = list(title = 'PC2'),
                                   zaxis = list(title = 'PC3')))
fig
```

# Contract data PCA analysis

2d PCA

```{r}
pca_df<-df%>%filter(group=="contract") %>% spread(mz,intensity)
pca_df$filename<-gsub(".*_", "", pca_df$filename)
pca_df$filename<-gsub("\\.txt$","",pca_df$filename)
pca_df$filename<-gsub(" ", "_", pca_df$filename)
pca_res <- prcomp(pca_df[,6:(ncol(pca_df)-1)], scale. = TRUE)

autoplot(pca_res, data = pca_df, colour = 'filename', size=2)
```

3d PCA

```{r}
dat_3d <- pca_res$x[,1:3] %>% as_tibble() %>% mutate(filenames = pca_df$filename)

fig <- plot_ly(dat_3d, x = ~PC1, y = ~PC2, z = ~PC3, 
               color =~filenames,colors=c('#8dc5f0','#43abf9','#0092ff','#cef37b','#a6d045','#7caa16','#f8bc7d','#e0892e','#ff8200'), size = I(150))
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'PC1'),
                                   yaxis = list(title = 'PC2'),
                                   zaxis = list(title = 'PC3')))

fig
```

# Correlation matrix

```{r}
source("http://www.sthda.com/upload/rquery_cormat.r")

contract<-df%>%filter(group=="contract")

contract$filename<-gsub(".*_", "", contract$filename)
contract$filename<-gsub("\\.txt$","",contract$filename)
contract$filename<-gsub(" ", "_", contract$filename)

contract%>%filter(group=="contract") %>% select("filename","mz","intensity") %>% spread(filename,intensity) %>% rquery.cormat()
```

CLEANING...

```{r}
rm(dat_3d,fig,pca_df,pca_res,pr.out)
```


# Histogram
First ahistogram is made to see the distribution of our data

```{r}

df$filename<-gsub("\\.txt$","",df$filename)
df %>% 
  select("filename","mz","intensity") %>% 
  group_by(intensity) %>%  
  summarize(count=n()) %>% 
  as.data.frame() %>% 
  ggplot(aes(x=intensity)) + geom_histogram(bins = 50, fill="dodgerblue", color ="black")
```

The coefficient of variation is used to select the most important ones

```{r}
df_cv<-df %>% 
  select("filename","mz","intensity") %>% 
  group_by(mz) %>%  
  summarize(CV=sd(intensity) / mean(intensity) * 100)

df_cv %>% 
  ggplot(aes(x=CV)) + geom_histogram(bins = 50, fill="dodgerblue", color ="black")

#Merging the dataframes 
df_hm <- merge(df_cv, df, by = "mz")
```

```{r}
#Trimming, taking the top 100 MZs. Since we have 50 samples, we have to take the top 5000 CVs.
df_hm %>% 
  top_n(5000,CV) %>% 
  select("filename","mz","intensity","CV") %>%
  group_by(mz) %>% 
  ggplot(aes(x=CV)) + geom_histogram(bins = 100, fill="dodgerblue", color ="black")
```

# PHEATMAP

<https://towardsdatascience.com/pheatmap-draws-pretty-heatmaps-483dab9a3cc>

```{r}
if("pheatmap" %in% rownames(installed.packages()) == FALSE) {install.packages("pheatmap")}
library(pheatmap)
```

## Top 100
```{r}
test<-df_hm %>% 
  top_n(5000,CV) %>% 
  unite("filename", pen:filename, remove=F) %>% 
  select("filename","mz","intensity")%>% 
  spread(filename,intensity)


test<-as.matrix(test[-1])

pheatmap(test,main = "pheatmap default", clustering_distance_cols = "correlation",clustering_distance_rows = "correlation")
```

## Top 50
```{r}
test<-df_hm %>% 
  top_n(2500,CV) %>% 
  unite("filename", pen:filename, remove=F) %>% 
  select("filename","mz","intensity")%>% 
  spread(filename,intensity)


test<-as.matrix(test[-1])

pheatmap(test,main = "pheatmap default", clustering_distance_cols = "correlation",clustering_distance_rows = "correlation")
```

## Top 25
```{r}
test<-df_hm %>% 
  top_n(1250,CV) %>% 
  unite("filename", pen:filename, remove=F) %>% 
  select("filename","mz","intensity")%>% 
  spread(filename,intensity)


test<-as.matrix(test[-1]) %>% t()

pheatmap(test,main = "pheatmap default", clustering_distance_cols = "correlation",clustering_distance_rows = "correlation")
```
